---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggpathdiagramr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggpathdiagramr)](https://CRAN.R-project.org/package=ggpathdiagramr)
<!-- badges: end -->

This package is in a very preliminary state. It has some preliminary plumbing but does not yet have its primary functions. Its future function will be to construct path diagrams in R via ggplot2, but with a TiKz-like workflow.

## Installation

You can install the development version of ggpathdiagramr like so:

``` r
remotes::install_github("wjschne/ggpathdiagramr")
```

## Things ggpathdiagramr can do so far

### Points

Points can be added and substracted

```{r points}
library(ggpathdiagramr)
p1 <- point(1,0)
p2 <- point(0,1)
p3 <- p1 + p2
p3
p3 - p2
```

Technically, a point is not a geometric vector, but the `point` object contains information about the vector that starts in the origin and ends at the point:

* `@distance`: The distance from the origin to the point (i.e., the vector's magnitude)
* `@slope`: The slope of the line containing the vector
* `@angle`: The angle (in radians) from the line on the x-axis to the line containing the vector.

```{r}
p3
p3@distance
p3@slope
p3@angle
```


A point can be converted to a matrix or a numeric vector:

```{r}
p3@xy
as.matrix(p3)
as.numeric(p3)
```

A list of points can be created like so:

```{r}
point(x = c(2,4,5), y = c(5,3,2))
```
A point can be created with radial coordinates

```{r}
point(distance = 5, angle = pi/ 2)
```


### Lines

Aside from a vertical line, lines can be constructed from a slope and an intercept

```{r}
line(slope = 1, intercept = 3)
```

Because the default slope is 0, a horizontal line can be set with just the intercept:

```{r}
line(intercept = 3)
```


A vertical line can be set 



Any line can be constructed from the coefficients of this equation:

$$
ax+by+c=0
$$



```{r}
line(a = 1, b = 2, c = 3)
```

### Segments

```{r}
s1 <- segment(p1 = point(0,1), p2 = point(3,3))
s1@line
```
