---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggpathdiagramr)
```

# ggpathdiagramr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggpathdiagramr)](https://CRAN.R-project.org/package=ggpathdiagramr)
<!-- badges: end -->

This package is in a very preliminary state. It has some structural beams and a bit of plumbing but is not yet capable of fulfilling its eventual purpose: To construct path diagrams in R via ggplot2, but with a TiKz-like workflow.

## Installation

You can install the development version of ggpathdiagramr like so:

``` r
remotes::install_github("wjschne/ggpathdiagramr")
```







### Lines



### Segments



### Retangles

A rectangle has 4 corners (`northeast`, `northwest`, `southwest`, and `southeast`). It has a center. It has width and height.
If you give the `rectangle` function enough information to calculate its four corners, it will do so. All of the following will give the same rectangle:


```{r}
ne <- point(4,2)
nw <- point(0,2)
sw <- point(0,0)
se <- point(4,0)
cent <- point(2,1)

# Give width, height, and any point
rectangle(width = 4, height = 2, center = cent)
rectangle(width = 4, height = 2, northeast = ne)
rectangle(width = 4, height = 2, northwest = nw)
rectangle(width = 4, height = 2, southwest = sw)
rectangle(width = 4, height = 2, southeast = se)

# Give the center and any of the 4 corners
rectangle(center = cent, northeast = ne)
rectangle(center = cent, northwest = nw)
rectangle(center = cent, southwest = sw)
rectangle(center = cent, southeast = se)

# Give opposite corners
rectangle(northeast = ne, southwest = sw)
rectangle(northwest = nw, southeast = se)

# Give width and two points either side
rectangle(width = 4, northwest = nw, southwest = sw)
rectangle(width = 4, northeast = ne, southeast = se)

# Give height and two points on top or bottom
rectangle(height = 2, northwest = nw, northeast = ne)
rectangle(height = 2, southwest = sw, southeast = se)
```


## Methods

### Distances

The distance function can be used to find the distance between two points.

```{r}
distance(p1, p2)
```

The distance of a point to the origin:

```{r}
distance(p1)
```


The distance between the endpoints of segment:

```{r}
distance(s1)
```

The distance from a point to a line:

```{r}
distance(p1, l1)
```


