---
title: "Angles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Angles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggpathdiagramr)
```

```{r setup}
library(ggpathdiagramr)
```

Angles have different kinds of units associated with them: degrees, radians, gradians, and turns.

I like $\pi$ just fine, but I agree with Michael Hartl's [Tau Manifesto](https://tauday.com/tau-manifesto) that we would have been better off if we had recognized that the number of radians to complete a full circle is more fundamental than the number of radians to complete a half circle. The symbol for this value is $\tau = 2\pi\approx 6.283185$

The `angle` class can take turns (i.e., 1 turn = one full rotation a circle), degrees (1 turn = 360 degrees), radians (1 turn = $2\pi$ = $\tau$), or gradians (1 turn = 400 gradians, gons, or grads).

One can create equivalent angles with any of the four metrics.

Turns | Radians | Degrees | Gradians|
:----:|:-------:|:-------:|:-------:|
.25 | .25&tau;<br>.5&pi; | 90&deg; | 100grad |

A right angle is
 * one quarter turn
 * 1 quarter of $\tau$ radians
 * one half of $\pi$ radians
 * 90 degrees
 * 100 gradians (AKA gons, grad, grades)

```{r}
tau <- 2 * pi
right_angle <- angle(turn = .25)
right_angle == angle(radian = .25 * tau)
right_angle == angle(radian = .5 * pi)
right_angle == angle(degree = 90)
right_angle == angle(gradian = 90)
```


# Arithmetic Operations

Angles can be added, subtracted, multiplied, and divided. The result is never negative and always less than 1 turn (360 degrees or $\tau$ radians).

$$30\unicode{xb0} + 60\unicode{xb0} = 90\unicode{xb0}$$

```{r}
angle(degree = 30) + angle(degree = 60)

```

$$\begin{align}330\unicode{xb0} + 60\unicode{xb0} &= 390\unicode{xb0}\\&=30\unicode{xb0}\end{align}$$

```{r}
angle(degree = 330) + angle(degree = 60)
```

$$\begin{align}10\unicode{xb0} - 20\unicode{xb0} &= -10\unicode{xb0}\\&=350\unicode{xb0}\end{align}$$

```{r}
angle(degree = 10) - angle(degree = 20)
```

$$2\cdot20\unicode{xb0}=40\unicode{xb0}$$


```{r}
2 * angle(degree = 20) 
```

$$\begin{align}2\cdot180 &= 360\unicode{xb0}\\&=0\unicode{xb0}\end{align}$$

```{r}
2 * angle(degree = 180) 
```

# Trigonometry

Angles can take the three standard trigonometric functions

```{r}
theta <- angle(degree = 60)
cos(theta)
sin(theta)
tan(theta)
```

Benefits of using trigonometric functions with angles instead of numeric radians include:

* Conversions from angles are handled automatically.
* Under the hood, the `cospi`, `sinpi`, and `tanpi` functions are used to get the rounding right on key locations (e.g., 90 degrees)


